---
author: "Alix F. Leon"
title: "Project 1: History of Philosophy"
subtitle: "A View at Philosophy Throughout the Years"
date: "Sep. 21, 2022"
output: html_document
---

```{r, echo = F, warning=F,message=F}
library(tidyverse)
library(ggplot2)
library(sentimentr)
```

```{r, echo = F, warning=F,message=F}
theme_set(theme_light())
data_path <- "../data/philosophy_data.csv"
```

```{r, echo = F, warning=F,message=F}
data <- read_csv(data_path)
data$school <- data$school %>% str_to_title() %>% str_replace("_", " ")
```
```{r, echo = F, warning=F,message=F}
head(data)
```

```{r}
data %>%
    ggplot(aes(x = school)) +
        geom_bar() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Philosophy has played a signifficant roll throughout human history. It is an
attempt at understanding knowledge, nature, reality, values, ethics, among
many other aspects of our lives through the use rational thought. Many of
whom we now consider some of the greatest minds have participated in 
philosophical reasoning, and published their ideas in an effort to immortalize
their findings. Consequently, many other authors have published a plethora of
philosophical ideas regarding the vast multitude of human concerns leading to
creation of several philosophical schools of thought.

This analysis will take at look at philosophical texts from 13 philosophical schools of thought. 
The data for this analysis was extracted from the [History of philosophy][dataset_link]
dataset. Although many sophisticated models can be used in conjunction with this data, we will use a 
more traditional approach, using statistical analyses and plots to observe trends, and ultimately
support the story.

### Dark ages
```{r, echo=F, message=F, fig.align='center'}
# Number of publications over the years by school
data %>%
    group_by(school, original_publication_date) %>%
    summarise(n_publications = n()) %>%
    mutate(school = as_factor(school)) %>%
    ggplot(
        aes(
            x = original_publication_date,
            y = log10(n_publications),
            color = school,
        )
    ) +
    geom_point() +
    labs(
        title = "Number of publications over the years by school",
        subtitle = str_wrap(
            "There is a gap in text published between
            the years 170 A.D. and 1637 A.D.",
            width = 70),
        x = "Original Publication Date",
        y = "Number of Publications (log10)",
        color = "School of Philosophy"
    ) +
    theme(
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom"
    )
```

```{r}
# data %>%
#     group_by(original_publication_date) %>%
#     summarise(n_publications = n()) %>%
#     ggplot(aes(x = original_publication_date, y = n_publications)) +
#         geom_bar()
```

```{r}
# Number of publications for each school
data %>%
    ggplot(aes(x = school)) +
        geom_bar(position = "identity") +
        theme(
            axis.text.x = element_text(
                angle = 45, hjust = 1
            )
        )
```


<!-- Links -->
[dataset_link]: https://www.kaggle.com/datasets/kouroshalizadeh/history-of-philosophy