---
author: "Alix F. Leon"
title: "Project 1: History of Philosophy"
subtitle: "A View at Philosophy Throughout the Years"
date: "Sep. 21, 2022"
output: html_document
---

```{r, echo = F, warning=F,message=F}
library(tidyverse)
library(ggplot2)
library(sentimentr)
```

```{r, echo = F, warning=F,message=F}
theme_set(theme_light())
data_path <- "../data/philosophy_data.csv"
```

```{r, echo = F, warning=F,message=F}
data <- read_csv(data_path)
data$school <- data$school %>% str_to_title() %>% str_replace("_", " ")
```
```{r, echo = F, warning=F, message=F, include=F}
head(data)
```

```{r, echo=F, warning=F, message=F}
data %>%
    ggplot(aes(x = school)) +
        geom_bar() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Philosophy has played a signifficant roll throughout human history. It is an
attempt at understanding knowledge, nature, reality, values, ethics, among
many other aspects of our lives through the use rational thought. Many of
whom we now consider some of the greatest minds have participated in 
philosophical reasoning, and published their ideas in an effort to immortalize
their findings. Consequently, many other authors have published a plethora of
philosophical ideas regarding the vast multitude of human concerns leading to
creation of several philosophical schools of thought.

This analysis will take at look at philosophical texts from 13 philosophical schools of thought. 
The data for this analysis was extracted from the [History of philosophy][dataset_link]
dataset. Although many sophisticated models can be used in conjunction with this data, we will use a 
more traditional approach, using statistical analyses and plots to observe trends, and ultimately
support the story.

## Dark ages
```{r, echo=F, message=F}
publications_by_school <- data %>%
    group_by(school, original_publication_date) %>%
    summarise(n_publications = n()) %>%
    mutate(school = as_factor(school)) %>%
    arrange(original_publication_date)
```

```{r, echo=F, message=F, fig.align='center', fig.width=10}
# Number of publications over the years by school
publications_by_school %>%
    ggplot(
        aes(
            x = original_publication_date,
            y = log10(n_publications),
            color = school,
        )
    ) +
    geom_point() +
    labs(
        title = "Number of publications over the years by school",
        subtitle = str_wrap(
            "There is a gap in text published between
            the years 170 A.D. and 1637 A.D.",
            width = 75
        ),
        x = "Original Publication Date",
        y = "Number of Publications (log10)",
        color = "School of Philosophy"
    ) +
    theme(
        legend.position = "bottom"
    )
```

Despite the influence of philosphy over the course of history, there have been
some "pathches" along the way. Although the data does not represent a complete picture
of philosophical texts published throughout the years, we can see from the plot above
that there have been several instances in which barely any texts have been published.
Take as an example the five oldest philosophical schools of thought in our data, i.e.,
the school of [Plato], the school of [Aristotle], the school of [stoicism], the school
of [Rationalism], and the school of [Empiricism]. 

```{r, echo = F, message=F, warning=F}
first_and_last_publication <- publications_by_school %>%
                                group_by(school) %>%
                                summarise(
                                    "min date" = min(original_publication_date),
                                    "max date" = max(original_publication_date)
                                ) %>%
                                arrange(`min date`)

difference_in_years <- tibble(
    schools = paste(
        first_and_last_publication$school,
        first_and_last_publication$school[2:13],
        sep = "-"
    ) %>% as_factor(),
    date_difference = (
        lead(first_and_last_publication["max date"])
        - first_and_last_publication["min date"]
    ) %>% unlist()
) %>%
drop_na(date_difference)
```
```{r, echo=F, message=F, warning=F, fig.align="center", fig.width=10}
difference_in_years %>%
    ggplot(aes(x = schools, y = date_difference)) +
        geom_bar(stat = "identity") +
        theme(
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom"
    ) +
    labs(
        title = "Time Elapsed Between Philosophical Schools of Thought",
        x = paste(
            "Contiguous Philosophical Schools",
            "of Thought (by Original Publication Date)"
        ),
        y = "Difference in Years"
    )
```
The chart above effectively shows that there is a margin of ~500 years between publication for
the school of Aristotle and the school of Stoicism, and more noticeably, a difference of more
than ~1,500 years between publications from the school of Stoicism and the school of Rationalism. 
This last time margin aligns with a time period in history denominated the [early middle ages].
During this time period, there was a signifficant reduction in the number of philosophical
publications[ยน][1].

## Philosophical Publications and Most Published Authors
```{r, echo=F, message=F, warning=F,}
# Add a flag for texts published before the middle ages
# Note: 500 was chosen arbitrarily as a nice division
# point for pre middle ages publications and post middle
# ages publications.
pre_middle_ages_publications <- publications_by_school %>%
    mutate(pre_middle_ages = original_publication_date <= 500)
```

```{r, echo=F, message=F, warning=F, fig.align="center", fig.width=10}

pre_middle_ages_publications %>%
    arrange(original_publication_date, n_publications) %>%
    ggplot(aes(x = school, y = n_publications)) +
        geom_col(
            aes(fill = as_factor(pre_middle_ages))
        ) +
        labs(
            title = "Number of Publications by School of Thought",
            subtitle = str_wrap(
                "Each philosophical school is colored
                according to whether a text was published
                before or after the middle ages",
                width = 75
            ),
            x = "Philosophical School of Thought",
            y = "Number of Publications",
            fill = "Before early ages"
        ) +
        theme(
            axis.text.x = element_text(
                angle = 45, hjust = 1
            )
        )
```

Regardless of period of declined philosophical activity, there have been many texts published before
and after this gap. For instance, the [Analytic] school and the school of [Aristotle] combined account
for more than 100 thousand philosophical publications from 


## Sentiment Analysis

```{r}
# Number of publications for each school
data %>%
    ggplot(aes(x = school)) +
        geom_bar(position = "identity") +
        theme(
            axis.text.x = element_text(
                angle = 45, hjust = 1
            )
        )
```

## Sources
[1]: #sources

<!-- Links -->
[dataset_link]: https://www.kaggle.com/datasets/kouroshalizadeh/history-of-philosophy
[Plato]: https://en.wikipedia.org/wiki/Platonic_Academy
[Aristotle]: https://en.wikipedia.org/wiki/Peripatetic_school
[Stoicism]: https://en.wikipedia.org/wiki/Stoicism
[Rationalism]: https://en.wikipedia.org/wiki/Rationalism
[Empiricism]: https://en.wikipedia.org/wiki/Empiricism
[Analytic]: https://en.wikipedia.org/wiki/Analytic_philosophy
[early middle ages]: https://en.wikipedia.org/wiki/Early_Middle_Ages

<!-- [A companion to philosophy in the Middle Ages. Blackwell Pub.](https://books.google.com/books?id=hRJT7O-uNooC&lpg=PA1&ots=Ddy7vORhhG&dq=middle%20ages%20philosophy&lr&pg=PA1#v=onepage&q=middle%20ages%20philosophy&f=false)  -->