---
author: "Alix F. Leon - afl2124"
title: "Project 1: History of Philosophy"
subtitle: "A View at Philosophy Throughout the Years"
date: "Sep. 21, 2022"
output: html_document
---

```{r, echo = F, warning=F,message=F}
library(tidyverse)
library(ggplot2)
library(treemapify)
library(ggplotify)
library(sentimentr)
# library(d3treeR)
```

```{r, echo = F, warning=F,message=F}
theme_set(theme_light())
data_path <- "../data/philosophy_data.csv"
```

```{r, echo = F, warning=F,message=F}
data <- read_csv(data_path)
data$school <- data$school %>% str_to_title() %>% str_replace("_", " ")
```
```{r, echo = F, warning=F, message=F, include=F}
head(data)
```

Philosophy has played a signifficant roll throughout human history. It is an
attempt at understanding knowledge, nature, reality, values, ethics, among
many other aspects of our lives through the use rational thought. Many of
whom we now consider some of the greatest minds have participated in 
philosophical reasoning, and published their ideas in an effort to immortalize
their findings. Consequently, many other authors have published a plethora of
philosophical ideas regarding the vast multitude of human concerns leading to
creation of several philosophical schools of thought.

This analysis will take at look at philosophical texts from 13 philosophical schools of thought. 
The data for this analysis was extracted from the [History of philosophy][dataset_link]
dataset. Although many sophisticated models can be used in conjunction with this data, we will use a 
more traditional approach, using statistical analyses and plots to observe trends, and ultimately
support the story.

## Dark ages
```{r, echo=F, message=F}
sentences_by_publication <- data %>%
    group_by(title, school, author, original_publication_date) %>%
    summarise(n_sentences = n()) %>%
    mutate(school = as_factor(school)) %>%
    arrange(original_publication_date)
```

```{r, echo=F, message=F, fig.align='center', fig.width=10}
# Number of publication sentences over the years by school
sentences_by_publication %>%
    ggplot(
        aes(
            x = original_publication_date,
            y = log10(n_sentences),
            color = school,
        )
    ) +
    geom_point() +
    labs(
        title = "Number of sentences in publications over the years by school",
        subtitle = str_wrap(
            "There is a gap in the number of
            texts published between the years
            170 A.D. and 1637 A.D.",
            width = 75
        ),
        x = "Original Publication Date",
        y = "Number of Sentences (log10)",
        color = "School of Philosophy"
    ) +
    theme(
        legend.position = "bottom"
    )
```

Although the data does not represent a complete picture
of philosophical texts published throughout the years, we can see from the plot above
that there have been several instances in which barely any texts have been published.
Despite the influence of philosphy over the course of history, there have been
some "pathches" along the way. 

```{r, echo = F, message=F, warning=F}
first_and_last_publication <- sentences_by_publication %>%
                                group_by(school) %>%
                                summarise(
                                    "min date" = min(original_publication_date),
                                    "max date" = max(original_publication_date)
                                ) %>%
                                arrange(`min date`)

difference_in_years <- tibble(
    schools = paste(
        first_and_last_publication$school,
        first_and_last_publication$school[2:13],
        sep = "-"
    ) %>% as_factor(),
    date_difference = (
        lead(first_and_last_publication["max date"])
        - first_and_last_publication["min date"]
    ) %>% unlist()
) %>%
drop_na(date_difference)
```
```{r, echo=F, message=F, warning=F, fig.align='center', fig.width=10}
difference_in_years %>%
    ggplot(aes(x = schools, y = date_difference)) +
        geom_bar(stat = "identity") +
        theme(
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom"
    ) +
    labs(
        title = "Time Elapsed Between Philosophical Schools of Thought",
        x = paste(
            "Contiguous Philosophical Schools",
            "of Thought (by Original Publication Date)"
        ),
        y = "Difference in Years"
    )
```
The chart above effectively shows that there is a margin of ~500 years between publication for
the school of Aristotle and the school of Stoicism, and more noticeably, a difference of more
than ~1,500 years between publications from the school of Stoicism and the school of Rationalism. 
This gap aligns with a time period in history called the [early middle ages].
During this time period, there was a signifficant reduction in the number of philosophical
publications[ยน][1].

## Sentence Length
```{r, echo=F, message=F, warning=F,}
# Add a flag for texts published before the middle ages
# Note: 500 was chosen arbitrarily as a nice division
# point for pre middle ages publications and post middle
# ages publications.
pre_middle_ages_publications <- sentences_by_publication %>%
    mutate(pre_middle_ages = original_publication_date <= 500)
```

```{r, echo=F, message=F, warning=F, fig.align='center', fig.width=10}

pre_middle_ages_publications %>%
    arrange(original_publication_date, n_sentences) %>%
    ggplot(aes(x = school, y = n_sentences)) +
        geom_col(
            aes(fill = as_factor(pre_middle_ages))
        ) +
        labs(
            title = "Number of Publications by School of Thought",
            subtitle = str_wrap(
                "Each philosophical school is colored
                according to whether a text was published
                before or after the middle ages",
                width = 75
            ),
            x = "Philosophical School of Thought",
            y = "Number of Publications",
            fill = "Before early ages"
        ) +
        theme(
            legend.position = "bottom",
            axis.text.x = element_text(
                angle = 45, hjust = 1
            )
        )
```

```{r, echo = F, message=F, warning=F, include = F}
# Number of publications by the school of Plato and the school of Aristotle
num_before_middle_ages <- pre_middle_ages_publications %>%
    ungroup() %>%
    dplyr::filter(pre_middle_ages) %>%
    dplyr::select(n_sentences) %>%
    sum()

num_after_middle_ages <- nrow(data) - num_before_middle_ages

num_before_middle_ages / nrow(data)
```
Philosophical texts, both during the modern age and during its beginnings, tend to be very dense.
For instance, the works of authors before the middle ages combined account for more than
89 thousand sentences. In other words, these publications from the 
aforementioned schools account for roughly 24.85% of the sentences contained in the dataset.

```{r, echo=F, message=F, warning=F, fig.align='center', fig.height = 7, fig.width=10}
sentences_by_publication %>%
    ungroup() %>%
    ggplot(
        mapping = aes(
                area = n_sentences,
                subgroup = school,
                subgroup2 = author,
                fill = school,

    )) +
        geom_treemap() +
        geom_treemap_subgroup2_text(aes(label = author)) +
        scale_x_continuous(expand = c(0, 0)) +
        scale_y_continuous(expand = c(0, 0)) +
        labs(
            title = str_wrap(
                        "Treemap of Number of Sentences per
                        Author (Grouped By Philosophical School of Thought)",
                        width = 75
                    ),
            subtitle = str_wrap(
                        "Larger region areas correspond
                        to a larger number of sentences",
                        width = 75
                    ),
            fill = "Philosophical School of Thought"
        ) +
        theme(
            legend.position = "bottom"
        )
```

This treemap shows how different authors compare in terms of sentence length. The regions are 
divided by their respective philosophical schools of thought

## Sentiment Analysis

These publications cover a wide variety of topics; consequently, 
```{r}
# Number of publications for each school
data %>%
    ggplot(aes(x = school)) +
        geom_bar(position = "identity") +
        theme(
            axis.text.x = element_text(
                angle = 45, hjust = 1
            )
        )
```

## Sources
[1]: #sources

<!-- Links -->
[dataset_link]: https://www.kaggle.com/datasets/kouroshalizadeh/history-of-philosophy
[school of Plato]: https://en.wikipedia.org/wiki/Platonic_Academy
[school of Aristotle]: https://en.wikipedia.org/wiki/Peripatetic_school
[Stoicism]: https://en.wikipedia.org/wiki/Stoicism
[Rationalism]: https://en.wikipedia.org/wiki/Rationalism
[Empiricism]: https://en.wikipedia.org/wiki/Empiricism
[Analytic]: https://en.wikipedia.org/wiki/Analytic_philosophy
[early middle ages]: https://en.wikipedia.org/wiki/Early_Middle_Ages

<!-- [A companion to philosophy in the Middle Ages. Blackwell Pub.](https://books.google.com/books?id=hRJT7O-uNooC&lpg=PA1&ots=Ddy7vORhhG&dq=middle%20ages%20philosophy&lr&pg=PA1#v=onepage&q=middle%20ages%20philosophy&f=false)  -->